Leadership across cultures and contexts: an exploratory analysis

### 2020-06-22 leadership_across_cultures_contexts.Rmd
Minor edits in conclusion.

Coercive authority is a dimension of leadership that is common in the ethnographic record and for which evidence does not vary by any of our four factors, yet did not quite make the 60% cutoff to be included on our list of candidate universals. This dimension features in several evolutionary models of leadership involving dominance hierarchies [@barkow_darwin_1989; @tiger_imperial_1971; @henrich_evolution_2001; @cheng_dominance_2019; @van_vugt_dual_2019] and the role of punishment in promoting cooperation [e.g., @marlowe_more_2008; @henrich_markets_2010; @henrich_costly_2006; @boyd_punishment_1992]. Indeed, we found that coercive authority among leaders was associated with enforcing punishment, aggressiveness, being feared, and also with controlling economics, which evokes "resource holding power", a concept from the behavioral ecology of dominance hierarchies (Figure \@ref(fig:elasticcoercive)). This passage describing the authority of Trobriand chiefs is illustrative [@oj29-007, pp. 64-65]:

@@

Relationships between leadership and high status are currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019; @garfield_evolutionary_2019]. Although leadership and high social status often covary, leaders are not always high-status individuals, nor are high status individuals always leaders [see, e.g., @van_vugt_dual_2019; @cheng_why_2020; @garfield_review_2019].

@@

# Concluding remarks: Beyond the non-WEIRD 
<!-- CAN'T JUST BE NONWEIRD -->
In recognizing the limited generalizability of most extant data [@henrich_weirdest_2010], many social scientists are now diversifying their samples to better understand *human* psychology, behavior, organizations, and culture. We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution. Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity among populations that are now often categorized as "non-WEIRD", "traditional", or "small-scale". At the same time, anthropologists have made the opposite mistake in often essentializing diverse "others" and failing to recognize deep similarities among peoples of all cultures [@brown_human_1991; @said1979; @reyes-garcia_small-scale_2017].

>>>>

Coercive authority is a common leadership dimension in the ethnographic record and for which evidence does not vary by any of our four factors, yet did not quite make the 60% cutoff to be included on our list of candidate universals. This dimension features in several evolutionary models of leadership involving dominance hierarchies [@barkow_darwin_1989; @tiger_imperial_1971; @henrich_evolution_2001; @cheng_dominance_2019; @van_vugt_dual_2019] and the role of punishment in promoting cooperation [e.g., @marlowe_more_2008; @henrich_markets_2010; @henrich_costly_2006; @boyd_punishment_1992]. Indeed, we found that coercive authority among leaders was associated with enforcing punishment, aggressiveness, being feared, and also with controlling economics, which evokes "resource holding power", a concept from the behavioral ecology of dominance hierarchies (Figure \@ref(fig:elasticcoercive)). This passage describing the authority of Trobriand chiefs is illustrative [@oj29-007, pp. 64-65]:

@@

Relationships between leadership and high status are currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019]. Although leadership and high social status often covary, leaders are not always high-status individuals, nor are high status individuals always leaders [see, e.g., @van_vugt_dual_2019; @cheng_why_2020; @garfield_review_2019].

@@

# Concluding remarks: Beyond the non-WEIRD 
<!-- CAN'T JUST BE NONWEIRD -->
In recognizing the limited generalizability of most extant data [@henrich_weirdest_2010], many social scientists are now diversifying their samples to better understand *human* psychology, behavior, organizations, and culture. We caution that the WEIRD people problem in behavioral science cannot be remedied simply by a "non-WEIRD" solution. Our results, and much of anthropology and ethnography, clearly demonstrate enormous diversity among populations that are now often categorized as "non-WEIRD", "traditional", or "small-scale." Anthropologists have often made the opposite mistake in essentializing diverse "others" and failing to recognize deep similarities among peoples of all cultures [@brown_human_1991; @said1979; @reyes-garcia_small-scale_2017].
